---
title: "RR steelhead 2007-2020, Inbreeding, Notebook 12"
output:
  html_document:
    df_print: paged
---

# looking at related parents in the Russian River steelhead

# get the data
```{r}
library(tidyverse)
library(lubridate)

spawner <- read_csv("RR_steelhead_07-20_known_sexes_corrected_lengths.csv")
trios <- read_csv("RR_2007-2020_pedigree_ages_lengths.csv", guess_max = 2e4) %>%
  mutate(parent_dayofyear = as.integer(parent_dayofyear))

spawner
```
```{r}
full_spawner_meta <- read_rds("RR_spawner_meta.rds")
full_spawner_meta

meta_we_want <- full_spawner_meta %>% select(NMFS_DNA_ID, SAMPLE_ID, SAMPLE_COMMENTS, Sample_ID) %>% rename(indiv = NMFS_DNA_ID)

spawners_w_meta <- left_join(spawner, meta_we_want, by = "indiv")
spawners_w_meta 
```
```{r}
kid_meta <- meta_we_want %>% select(-Sample_ID) %>% rename(kid = indiv, kid_sample_id = SAMPLE_ID, kid_comments = SAMPLE_COMMENTS)
ma_meta <- meta_we_want %>% select(-Sample_ID) %>% rename(ma = indiv, ma_sample_id = SAMPLE_ID, ma_comments = SAMPLE_COMMENTS)
pa_meta <- meta_we_want %>% select(-Sample_ID) %>% rename(pa = indiv, pa_sample_id = SAMPLE_ID, pa_comments = SAMPLE_COMMENTS)

trio_meta <- trios %>% left_join(kid_meta, by = "kid") %>% 
  left_join(ma_meta, by = "ma") %>% 
  left_join(pa_meta, by = "pa")

trio_meta
```



# Looking at inbreeding in the RR
## How often are the parents half sibs/full sibs? How often are the parents half first cousins/full first cousins?
```{r}
#want to join the data about maternal and paternal grandparents. this will work for all the trios in teh age matrix whwere i have parents (and ages) for all the kids and the ma and pa
mat_gdparents <- trio_meta %>% select(kid, ma, pa) %>% rename(ma = kid, maternal_grandma = ma, maternal_grandpa = pa)
pat_gdparents <- trio_meta %>% select(kid, ma, pa) %>% rename(pa = kid, paternal_grandma = ma, paternal_grandpa = pa)

mat_great_gparents1 <- trio_meta %>% select(kid, ma, pa) %>% rename(maternal_grandma = kid, maternal_grandma_ma = ma, maternal_grandma_pa = pa)
mat_great_gparents2 <- trio_meta %>% select(kid, ma, pa) %>% rename(maternal_grandpa = kid, maternal_grandpa_ma = ma, maternal_grandpa_pa = pa)

#maternal grandma's ma == maternal_grandma_ma

pat_great_gparents1 <- trio_meta %>% select(kid, ma, pa) %>% rename(paternal_grandma = kid, paternal_grandma_ma = ma, paternal_grandma_pa = pa)
pat_great_gparents2 <- trio_meta %>% select(kid, ma, pa) %>% rename(paternal_grandpa = kid, paternal_grandpa_ma = ma, paternal_grandpa_pa = pa)


trio_data <- trio_meta %>% 
  mutate(ppair = paste(ma, pa, sep = "_")) %>% 
  filter(!is.na(ma_age), !is.na(pa_age)) %>% 
  distinct(ppair, .keep_all = T) %>%
  left_join(mat_gdparents, by = "ma") %>% 
  left_join(pat_gdparents, by = "pa") %>% 
  left_join(mat_great_gparents1, by = "maternal_grandma") %>% 
  left_join(mat_great_gparents2, by = "maternal_grandpa") %>% 
  left_join(pat_great_gparents1, by = "paternal_grandma") %>% 
  left_join(pat_great_gparents2, by = "paternal_grandpa") %>% 
  mutate(ma_pa_relationship = case_when((maternal_grandma == paternal_grandma & maternal_grandpa == paternal_grandpa) ~ "full_sib",
                                        maternal_grandma == paternal_grandma ~ "half_sib",
                                        maternal_grandpa == paternal_grandpa ~ "half_sib",
                                        (maternal_grandma_ma == paternal_grandma_ma & maternal_grandma_pa == paternal_grandma_pa) ~ "first_cousin",
                                        (maternal_grandpa_ma == paternal_grandpa_ma & maternal_grandpa_pa == paternal_grandpa_pa) ~ "first_cousin",
                                        (maternal_grandma_ma == paternal_grandpa_ma & maternal_grandma_pa == paternal_grandpa_pa) ~ "first_cousin",
                                        (maternal_grandpa_ma == paternal_grandma_ma & maternal_grandpa_pa == paternal_grandma_pa) ~ "first_cousin",
                                        maternal_grandma_ma == paternal_grandma_ma ~ "half_first_cousin",
                                        maternal_grandma_ma == paternal_grandpa_ma ~ "half_first_cousin",
                                        maternal_grandma_pa == paternal_grandma_pa ~ "half_first_cousin",
                                        maternal_grandma_pa == paternal_grandpa_pa ~ "half_first_cousin",
                                        maternal_grandpa_ma == paternal_grandma_ma ~ "half_first_cousin",
                                        maternal_grandpa_ma == paternal_grandpa_ma ~ "half_first_cousin",
                                        maternal_grandpa_pa == paternal_grandma_pa ~ "half_first_cousin",
                                        maternal_grandpa_pa == paternal_grandpa_pa ~ "half_first_cousin",
                                        TRUE ~ "other"))
         

#3 full sibs, 32 half sibs that are ma and pa, out of 3807 unique parent pairs
3/(3772+32+3)
32/(3772+32+3)

```


# Now let's find the recorded mating pairs that did not have any offspring return as adults (2013-2020)
```{r}
try_it_all <- trio_meta %>% 
  select(kid, ma, pa, ma_sample_id, pa_sample_id, everything()) %>%
  filter(SpawnYear >= 2013) %>% 
  group_by(SpawnYear, ma_hatchery, parent_dayofyear) %>% 
  mutate(ma_id = str_sub(ma_sample_id, -4,-1)) %>% 
  mutate(pa_id = str_sub(pa_sample_id, -5,-1)) %>% 
  mutate(ma_id_2 = str_sub(ma_id, 1,-2)) %>% 
  mutate(pa_id_2 = str_sub(pa_id, 1,-3)) %>% 
  mutate(pa_id_n = str_sub(pa_id, -1,-1)) %>% 
  select(ma_sample_id, ma_id, ma_id_2, pa_sample_id, pa_id, pa_id_2, pa_id_n, everything()) %>% 
  filter(ma_id_2 == pa_id_2) %>% 
  arrange(ma_sample_id, pa_id_n) %>% 
  mutate(pa_id_n = replace(pa_id_n, pa_id_n == "B", "b")) %>% 
  mutate(pa_id_n = replace(pa_id_n, pa_id_n == "C", "c")) %>%
  ungroup() %>% 
  group_by(SpawnYear, ma_hatchery, parent_dayofyear, ma_id_2) %>% 
  count(pa_id_n) %>% 
#write_csv(., "RR_inbreeding_finding_missing_pairs.csv")
  pivot_wider(., names_from = pa_id_n, values_from = n)
  
try_it_all %>% ungroup %>% count(SpawnYear)

missing_all <- try_it_all %>% filter(is.na(a) | is.na(b) | is.na(c)) %>% 
  mutate(missing = case_when((is.na(a) & is.na(b)) ~ "Ma,Mb",
                             (is.na(a) & is.na(c)) ~ "Ma,Mc",
                             (is.na(b) & is.na(c)) ~ "Mb,Mc",
                             is.na(a) ~ "Ma",
                             is.na(b) ~ "Mb",
                             is.na(c) ~ "Mc")) %>% 
  separate(missing, into = c("missing_1", "missing_2"), sep = ",") %>% 
  pivot_longer(cols = missing_1:missing_2, values_to = "missing") %>% 
  select(ma_id_2, missing) %>% 
  filter(!is.na(missing)) %>% 
  mutate(missing_males = paste(ma_id_2, missing, sep = "")) %>% 
  ungroup()

missing_all %>% ungroup() %>% count(SpawnYear)

ids_split <- trio_meta %>% 
  select(kid, ma, pa, ma_sample_id, pa_sample_id, everything()) %>%
  filter(SpawnYear >= 2013) %>% 
  group_by(SpawnYear, ma_hatchery, parent_dayofyear) %>% 
  mutate(ma_id = str_sub(ma_sample_id, -4,-1)) %>% 
  mutate(pa_id = str_sub(pa_sample_id, -5,-1)) %>% 
  mutate(ma_id_2 = str_sub(ma_id, 1,-2)) %>% 
  mutate(pa_id_2 = str_sub(pa_id, 1,-3)) %>% 
  mutate(pa_id_n = str_sub(pa_id, -1,-1)) %>% 
  select(ma_sample_id, ma_id, ma_id_2, pa_sample_id, pa_id, pa_id_2, pa_id_n, everything()) %>% 
  group_by(ma_hatchery, parent_dayofyear) %>% 
  filter(ma_id_2 == pa_id_2) %>% 
  arrange(ma_sample_id, pa_id_n) %>% 
  mutate(pa_id_n = replace(pa_id_n, pa_id_n == "B", "b")) %>% 
  mutate(pa_id_n = replace(pa_id_n, pa_id_n == "C", "c")) %>% 
  ungroup()

missing_pairs <- missing_all %>% 
  left_join(ids_split, by = c("SpawnYear","ma_hatchery", "parent_dayofyear", "ma_id_2")) %>% 
  select(ma_sample_id, missing_males, SpawnYear, ma_hatchery) %>% 
  mutate(missing_male_id = paste(str_sub(ma_sample_id, 1,-5), missing_males, sep = "")) %>% 
  ungroup() %>% 
  select(ma_sample_id, missing_male_id, SpawnYear, ma_hatchery) %>% 
  mutate(pair = paste(ma_sample_id, missing_male_id, sep = "_")) %>% 
  distinct(pair, .keep_all = T) %>% 
  mutate(idx = row_number())

missing_pairs %>% count(SpawnYear)
missing_pairs
```

```{r}
#maternal parents of missing pair
mat_parents_mp <- missing_pairs %>% 
  select(ma_sample_id, idx) %>% 
  left_join(trio_meta, by = c("ma_sample_id" = "kid_sample_id")) %>% 
  filter(!is.na(kid)) %>% 
  select(ma_sample_id, kid, ma, pa, idx) %>% 
  rename(f_sample_id = ma_sample_id, female = kid, ma_f = ma, pa_f = pa)

pat_parents_mp <- missing_pairs %>% 
  select(missing_male_id, idx) %>% 
  left_join(trio_meta, by = c("missing_male_id" = "kid_sample_id")) %>% 
  filter(!is.na(kid)) %>% 
  select(missing_male_id, kid, ma, pa, idx) %>% 
  rename(m_sample_id = missing_male_id, male = kid, ma_m = ma, pa_m = pa)

gparents1_mp <- mat_parents_mp %>% 
  select(ma_f, idx) %>% 
  left_join(trio_meta, by = c("ma_f" = "kid")) %>% 
  select(ma_f, ma, pa) %>% 
  rename(mat_gma_f = ma, mat_gpa_f = pa) %>% 
  distinct(ma_f, .keep_all = T)

#ma_f = mother of the female
#mat_gma_f = maternal grandma of the female

gparents2_mp <- mat_parents_mp %>% 
  select(pa_f, idx) %>% 
  left_join(trio_meta, by = c("pa_f" = "kid")) %>% 
  select(pa_f, ma, pa) %>% 
  rename(pat_gma_f = ma, pat_gpa_f = pa) %>% 
  distinct(pa_f, .keep_all = T)

gparents3_mp <- pat_parents_mp %>% 
  select(ma_m, idx) %>% 
  left_join(trio_meta, by = c("ma_m" = "kid")) %>% 
  select(ma_m, ma, pa) %>% 
  rename(mat_gma_m = ma, mat_gpa_m = pa) %>% 
  distinct(ma_m, .keep_all = T)

gparents4_mp <- pat_parents_mp %>% 
  select(pa_m, idx) %>% 
  left_join(trio_meta, by = c("pa_m" = "kid")) %>% 
  select(pa_m, ma, pa) %>% 
  rename(pat_gma_m = ma, pat_gpa_m = pa) %>% 
  distinct(pa_m, .keep_all = T)
```

# Now lets find the related spawn pairs out of the group that did not produce offspring 
```{r}
#no_offspring <- mat_parents_mp %>% 
#  inner_join(pat_parents_mp, by = c("idx")) %>% 
#  mutate(f_m_relationship = case_when((ma_f == ma_m & pa_f == pa_m) ~ "full_sib",
 #                                       ma_f == ma_m ~ "half_sib",
  #                                      pa_f == pa_m ~ "half_sib")) 

no_offspring  <- missing_pairs %>% 
  mutate(ppair = paste(ma_sample_id, missing_male_id, sep = "_")) %>% 
  distinct(ppair, .keep_all = T) %>%
  rename(f_sample_id = ma_sample_id,  m_sample_id = missing_male_id ) %>% 
  inner_join(mat_parents_mp, by = c("f_sample_id", "idx")) %>% 
  inner_join(pat_parents_mp, by = c("m_sample_id", "idx")) %>% 
  inner_join(gparents1_mp, by = "ma_f") %>% 
  inner_join(gparents2_mp, by = "pa_f") %>% 
  inner_join(gparents3_mp, by = "ma_m") %>% 
  inner_join(gparents4_mp, by = "pa_m") %>% 
  mutate(f_m_relationship = case_when((ma_f == ma_m & pa_f == pa_m) ~ "full_sib",
                                        ma_f == ma_m ~ "half_sib",
                                        pa_f == pa_m ~ "half_sib",
                                        (mat_gma_f == mat_gma_m & mat_gpa_f == mat_gpa_m) ~ "first_cousin",
                                        (mat_gma_f == pat_gma_m & mat_gpa_f == pat_gpa_m) ~ "first_cousin",
                                        (pat_gma_f == mat_gma_m & pat_gpa_f == mat_gpa_m) ~ "first_cousin",
                                        (pat_gma_f == pat_gma_m & pat_gpa_f == pat_gpa_m) ~ "first_cousin",
                                         mat_gma_f == mat_gma_m ~ "half_first_cousin",
                                         mat_gpa_f == mat_gpa_m ~ "half_first_cousin",
                                         mat_gma_f == pat_gma_m ~ "half_first_cousin",
                                         mat_gpa_f == pat_gpa_m ~ "half_first_cousin",
                                         pat_gma_f == mat_gma_m ~ "half_first_cousin",
                                         pat_gpa_f == mat_gpa_m~ "half_first_cousin",
                                         pat_gma_f == pat_gma_m~ "half_first_cousin",
                                         pat_gpa_f == pat_gpa_m~ "half_first_cousin",
                                        TRUE ~ "other"))
           



no_offspring

no_offspring %>% 
  count(f_m_relationship)
no_offspring %>% count(SpawnYear)

```

```{r}
offspring_nums <- read_csv("RR_offspring_num_per_pair.csv") %>% separate(., id, sep = ":", into = c("pa", "ma")) %>% 
  select(-type) %>% 
  rename(n_offspring = n)
offspring_nums

trios_offspring_n <- trio_data %>%
  left_join(offspring_nums, by = c("ma", "pa"))

trios_offspring_n %>% 
  ggplot()+
  geom_histogram(aes(x= n_offspring, y =..density..))+
  facet_wrap(~ma_pa_relationship)

trios_offspring_n %>%
  group_by(ma_pa_relationship) %>% 
  summarize(mean(n_offspring, na.rm = T))
```

```{r}
trios_offspring_n %>% 
  filter(ma_pa_relationship == "full_sib") 

no_offspring %>% 
  filter(f_m_relationship == "full_sib")


```

#add ages for the female and male in the "no_offspring" dataset and see if any patterns emerge
```{r}
no_offspring
trios
kid_ages_only <- trios %>% 
  select(kid, kid_age)
no_offspring_age <- no_offspring %>% 
  left_join(kid_ages_only, by = c("female" = "kid")) %>% 
  rename(female_age = kid_age) %>% 
  left_join(kid_ages_only, by = c("male" = "kid")) %>% 
  rename(male_age = kid_age) %>% 
  mutate(pair_age = paste(male_age, female_age, sep = "_")) %>% 
  mutate(mid_age = ((female_age + male_age) / 2) )
           
  
none_midage <- no_offspring_age %>% 
  filter(SpawnYear >= 2013, SpawnYear < 2018) %>% 
  count(mid_age) %>% 
  mutate(freq = n/(sum(n)))

none_female <- no_offspring_age %>% 
  filter(SpawnYear >= 2013, SpawnYear < 2018) %>%
  count(female_age) %>% 
  mutate(freq = n/(sum(n)))

none_male <- no_offspring_age %>% 
  filter(SpawnYear >= 2013, SpawnYear < 2018) %>%
  count(male_age) %>% 
  mutate(freq = n/(sum(n)))

#compare the freq of age groups with no offspring to the frea of age groups of fish with offspring
offspring_ma <- trios %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_ma_sex == "Female") %>% 
  mutate(pair = paste(ma,pa , sep = "_")) %>% 
  filter(!is.na(ma_age)) %>% 
  distinct(ma, .keep_all = T) %>% 
  count(ma_age) %>% 
  mutate(freq = n/(sum(n)))

offspring_pa <- trios %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_pa_sex == "Male") %>% 
  mutate(pair = paste(ma,pa , sep = "_")) %>% 
  filter(!is.na(pa_age)) %>% 
  distinct(pa, .keep_all = T) %>% 
  count(pa_age) %>% 
  mutate(freq = n/(sum(n)))

offspring_pair <- trios %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_ma_sex == "Female", new_pa_sex == "Male") %>% 
  mutate(pair = paste(ma,pa , sep = "_")) %>% 
  mutate(pair_age = paste(ma_age, pa_age, sep = "_")) %>% 
  mutate(mid_age = ((ma_age + pa_age) / 2) ) %>% 
  filter(!is.na(mid_age)) %>% 
  distinct(pair, .keep_all = T) %>% 
  count(mid_age) %>% 
  mutate(freq = n/(sum(n)))
  
#no check if significantly different with chi squared test

chisq.test(x = none_female$n, p = offspring_ma$freq) #sig
chisq.test(x = none_male$n, p = offspring_pa$freq) #not sig
chisq.test(x= none_midage$n, p = offspring_pair$freq) #sig
```

# Join the two datasets

## filter for only 2013-2017
```{r}
this <- no_offspring %>% 
  mutate(n_offspring = 0) %>% 
  select(female, male, f_m_relationship, n_offspring, SpawnYear, ma_hatchery) %>% 
  rename(ma = female, pa = male, ma_pa_relationship = f_m_relationship, year = SpawnYear, hatchery = ma_hatchery) %>% 
  filter(year >= 2013, year <=2017)%>% 
  filter(ma != "M082455") %>% 
  mutate(ma_pa = str_c(ma, pa, sep = "_")) %>% 
  filter(ma_pa != "M092804_M092806" )
#getting rid of duplicates parent pairs -- havea  few parents with zero offpsring in this list, but hte next list has them haveing offspring

this %>% count(year)

that <- trios_offspring_n %>% 
  select(ma, pa, ma_pa_relationship, n_offspring, SpawnYear, ma_hatchery.x, ma_age, pa_age) %>% 
  rename(year = SpawnYear, hatchery = ma_hatchery.x) %>% 
  filter(year >= 2013, year <=2017)
that %>% count(year)

this_that <- bind_rows(this, that)
this_that %>% count(year)
this_that %>% count(hatchery)

#write_csv(this_that, "RR_2013_2017_parent_pairs_and_num_offspring.csv")

this_that %>% count(ma_age)
this_that %>% count(pa_age)
```

```{r}
this_that %>% 
  group_by(ma_pa_relationship) %>% 
  summarize(mean(n_offspring))

group_order <- c("full_sib", "half_sib", "first_cousin", "half_first_cousin", "other" )

inbreeding <- this_that %>% 
  group_by(ma_pa_relationship) %>% 
  summarize(mean(n_offspring)) %>% 
  rename(mean_offspring = "mean(n_offspring)") %>% 
  ggplot()+
  geom_col(aes(x = ma_pa_relationship, y = mean_offspring, fill = ma_pa_relationship))+
  scale_fill_manual(values = c("#9a532b","#503143","#c49b60", "#79ad9f", "#193439"))+
  scale_x_discrete(limits = group_order)+
  theme_classic()+
  scale_y_continuous(breaks = seq(0,1.5, 0.25))+
  theme(legend.position = "none")+
  labs(x = "Parent Pair Relationship", y = "Mean Offspring")
inbreeding

ggsave("inbreeding.png", plot = inbreeding, width = 5, height = 4)

this_that %>% 
  group_by(ma_pa_relationship) %>% 
  count()

this_that %>% 
  group_by(ma_pa_relationship) %>% 
  count(n_offspring != 0)

this_that %>% 
  count(ma_pa_relationship)
```

Raw numbers of related parent pairs (2013-2017):
14 full sibling pairs (2 pairs had offspring) = 2/14 = 14.3%
	Mean num offspring 0.3571429		
30 half sibling pairs (19 pairs had offspring) = 19/30 = 63.3%
	Mean num offspring 0.8333333	
25 first cousin pairs (14 pairs had offspring) = 14/25 = 56%
	Mean num offspring 0.8800000
59 half first cousin pairs (38 pairs had offspring) = 38/59 = 64.4%
	Mean num offspring 1.1525424	
3708 other pairs (2620 pairs had offspring) =2620/3708 = 70.66% 
	Mean num offspring 	1.524811		


