---
title: "RR Steelhead, Animal Model Results, Notebook 10"
output:
  html_notebook: default
  pdf_document: default
---

Animal model for age at maturity

```{r}
library(tidyverse)
library(MCMCglmm)
library(naniar)
```
The animal model was done on a computing cluster. The R script to run the model is "age_animal_model_genetic_corr.R" and the model was run six times as an array, following the script "age_animal_model_genetic_corr_array.sh"

# Run 1
```{r}
# Run 1
gen_corr_1 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_01.rds")

summary(gen_corr_1)

```

```{r}
plot(gen_corr_1$Sol)
plot(gen_corr_1$VCV)

autocorr(gen_corr_1$Sol)
autocorr.diag(gen_corr_1$Sol)

posterior.heritability1.4.FEM <- gen_corr_1$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_1$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_1$VCV[, "ma"] +
    gen_corr_1$VCV[, "SpawnYear"] + gen_corr_1$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_1$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_1$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_1$VCV[, "ma"] +
    gen_corr_1$VCV[, "SpawnYear"] + gen_corr_1$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_1$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_1$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_1$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)
```
# Run 2
```{r}
# Run 2
gen_corr_2 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_02.rds")

summary(gen_corr_2)
```
```{r}
plot(gen_corr_2$Sol)
plot(gen_corr_2$VCV)


posterior.heritability1.4.FEM <- gen_corr_2$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_2$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_2$VCV[, "ma"] +
    gen_corr_2$VCV[, "SpawnYear"] + gen_corr_2$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_2$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_2$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_2$VCV[, "ma"] +
    gen_corr_2$VCV[, "SpawnYear"] + gen_corr_2$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_2$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_2$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_2$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)
```
# Run 3
```{r}
# run 3
gen_corr_3 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_03.rds")

summary(gen_corr_3)
```
```{r}
plot(gen_corr_3$Sol)
plot(gen_corr_3$VCV)


posterior.heritability1.4.FEM <- gen_corr_3$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_3$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_3$VCV[, "ma"] +
    gen_corr_3$VCV[, "SpawnYear"] + gen_corr_3$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_3$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_3$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_3$VCV[, "ma"] +
    gen_corr_3$VCV[, "SpawnYear"] + gen_corr_3$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_3$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_3$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_3$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)
```
# Run 4
```{r}
gen_corr_4 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_04.rds")

summary(gen_corr_4)
```
```{r}
plot(gen_corr_4$Sol)
plot(gen_corr_4$VCV)


posterior.heritability1.4.FEM <- gen_corr_4$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_4$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_4$VCV[, "ma"] +
    gen_corr_4$VCV[, "SpawnYear"] + gen_corr_4$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_4$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_4$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_4$VCV[, "ma"] +
    gen_corr_4$VCV[, "SpawnYear"] + gen_corr_4$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_4$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_4$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_4$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)
```
# Run 5
```{r}
gen_corr_5 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_05.rds")

summary(gen_corr_5)
```
```{r}
plot(gen_corr_5$Sol)
plot(gen_corr_5$VCV)


posterior.heritability1.4.FEM <- gen_corr_5$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_5$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_5$VCV[, "ma"] +
    gen_corr_5$VCV[, "SpawnYear"] + gen_corr_5$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_5$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_5$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_5$VCV[, "ma"] +
    gen_corr_5$VCV[, "SpawnYear"] + gen_corr_5$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_5$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_5$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_5$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)
```
# Run 6

```{r}
gen_corr_6 <- read_rds("animal_model/outputs/genetic_corr/RR_genetic_corr_06.rds")

summary(gen_corr_6)
```

```{r}
plot(gen_corr_6$Sol)
plot(gen_corr_6$VCV)


posterior.heritability1.4.FEM <- gen_corr_6$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (gen_corr_6$VCV[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + gen_corr_6$VCV[, "ma"] +
    gen_corr_6$VCV[, "SpawnYear"] + gen_corr_6$VCV[, "new_kid_sexFemale.units"])
posterior.heritability1.4.MAL <- gen_corr_6$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (gen_corr_6$VCV[, "new_kid_sexMale:new_kid_sexMale.animal"] + gen_corr_6$VCV[, "ma"] +
    gen_corr_6$VCV[, "SpawnYear"] + gen_corr_6$VCV[, "new_kid_sexMale.units"])

mean(posterior.heritability1.4.FEM)
mean(posterior.heritability1.4.MAL)

corr.gen <- gen_corr_6$VCV[,
'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(gen_corr_6$VCV[,'new_kid_sexFemale:new_kid_sexFemale.animal']*gen_corr_6$VCV[,'new_kid_sexMale:new_kid_sexMale.animal'])

mean(corr.gen)
HPDinterval(corr.gen)

str(corr.gen)
```
# Can I join the data from the 6 runs and look for convergence between the different chains?
```{r}
# https://github.com/tmalsburg/MCMCglmm-intro

library(parallel)

all6_sol <- list(gen_corr_1$Sol, gen_corr_2$Sol, gen_corr_3$Sol, gen_corr_4$Sol, gen_corr_5$Sol, gen_corr_6$Sol)
all6_sol <- do.call(mcmc.list, all6_sol)

all6_vcv <- list(gen_corr_1$VCV, gen_corr_2$VCV, gen_corr_3$VCV, gen_corr_4$VCV, gen_corr_5$VCV, gen_corr_6$VCV)
all6_vcv <- do.call(mcmc.list, all6_vcv)

library(coda)

par(mfrow=c(4,2), mar=c(2,2,1,2))
gelman.plot(all6_sol, auto.layout=F)

par(mfrow=c(8,2), mar=c(2, 1, 1, 1))
plot(all6_sol, ask=F, auto.layout=F)
##
par(mfrow=c(4,2), mar=c(2,2,1,2))
gelman.plot(all6_vcv, auto.layout=F)

par(mfrow=c(8,2), mar=c(2, 1, 1, 1))
plot(all6_vcv, ask=F, auto.layout=F)

##
# test for convergence, Gelman-Runin, its a ratio, want it near 1, maybe up to 1.1 or 1.2
# https://bookdown.org/rdpeng/advstatcomp/monitoring-convergence.html#gelman-rubin-statistic

gelman.diag(all6_sol)


```
```{r}
# autocorrelation
# "The plots below show for each parameter the autocorrelation as a function of the distance between samples. If the distance is 0, the autocorrelation is one because the correlation of a variable with itself is one. However, as the distance between samples increases the autocorrelation diminishes. We also see that the parameter for the intercept has much lower autocorrelation than the other parameters." - https://github.com/tmalsburg/MCMCglmm-intro

plot.acfs <- function(x) {
  n <- dim(x)[2]
  par(mfrow=c(ceiling(n/2),2), mar=c(3,2,3,0))
  for (i in 1:n) {
    acf(x[,i], lag.max=100, main=colnames(x)[i])
    grid()
  }
}
plot.acfs(gen_corr_1$Sol)
plot.acfs(gen_corr_1$VCV)
```
```{r}
# need to get the model runs together to calculate the gentic correlations 
# gen_corr_1$Sol, gen_corr_2$Sol, gen_corr_3$Sol, gen_corr_4$Sol, gen_corr_5$Sol, gen_corr_6$Sol

tib_1 <- as_tibble(gen_corr_1$Sol)
tib_2 <- as_tibble(gen_corr_2$Sol) 
tib_3 <- as_tibble(gen_corr_3$Sol)
tib_4 <- as_tibble(gen_corr_4$Sol)
tib_5 <- as_tibble(gen_corr_5$Sol)
tib_6 <- as_tibble(gen_corr_6$Sol)

all_sol_tib <- bind_rows(tib_1,tib_2, tib_3, tib_4, tib_5, tib_6)

vcv_list <- list(
  tib_1 = as_tibble(gen_corr_1$VCV),
  tib_2 = as_tibble(gen_corr_2$VCV), 
  tib_3 = as_tibble(gen_corr_3$VCV),
  tib_4 = as_tibble(gen_corr_4$VCV),
  tib_5 = as_tibble(gen_corr_5$VCV),
  tib_6 = as_tibble(gen_corr_6$VCV)
)

all_vcv_tib <- bind_rows(vcv_list, .id = "MCMC_Run")

all_sol_tib
all_vcv_tib
```



```{r}
##
#genetic correlation from the models grouped together?

corr.gen.all <-  (all_vcv_tib[,'new_kid_sexFemale:new_kid_sexMale.animal']/sqrt(all_vcv_tib[,'new_kid_sexFemale:new_kid_sexFemale.animal']*all_vcv_tib[,'new_kid_sexMale:new_kid_sexMale.animal']))

mean(corr.gen.all$'new_kid_sexFemale:new_kid_sexMale.animal')

library(HDInterval)

hdi(corr.gen.all)


corr.gen.all %>% 
  rename(f_m = 'new_kid_sexFemale:new_kid_sexMale.animal' ) %>% 
  ggplot()+
  geom_density(aes(x = f_m ))
```

```{r}
#heritability from all the models grouped together

herit_female_all <- all_vcv_tib[, "new_kid_sexFemale:new_kid_sexFemale.animal"] /
  (all_vcv_tib[, "new_kid_sexFemale:new_kid_sexFemale.animal"] + all_vcv_tib[, "ma"] +
    all_vcv_tib[, "SpawnYear"] + all_vcv_tib[, "new_kid_sexFemale.units"])

herit_male_all <- all_vcv_tib[, "new_kid_sexMale:new_kid_sexMale.animal"] /
  (all_vcv_tib[, "new_kid_sexMale:new_kid_sexMale.animal"] + all_vcv_tib[, "ma"] +
    all_vcv_tib[, "SpawnYear"] + all_vcv_tib[, "new_kid_sexMale.units"])

mean(herit_female_all$'new_kid_sexFemale:new_kid_sexFemale.animal')
hdi(herit_female_all)

herit_female_all %>% 
  rename(f_f = 'new_kid_sexFemale:new_kid_sexFemale.animal' ) %>% 
  ggplot()+
  geom_density(aes(x = f_f ))

mean(herit_male_all$'new_kid_sexMale:new_kid_sexMale.animal')
hdi(herit_male_all)

herit_male_all %>% 
  rename(m_m = 'new_kid_sexMale:new_kid_sexMale.animal' ) %>% 
  ggplot()+
  geom_density(aes(x = m_m ))
```
effective sample size
```{r}
effectiveSize(herit_male_all)
effectiveSize(herit_female_all)
effectiveSize(corr.gen.all)
```


```{r}
all_sol_tib
all_vcv_tib

all6_sol %>% summary()
all6_vcv %>% summary()
```

