---
title: "RR Steelhead 2007-2020, Notebook 1"
subtitle: "Cleaning up the genotype data"
output: html_notebook
---

Uploading the data.
```{r}
library(tidyverse)
library(lubridate)
RRSteelhead <- read_csv( "RR_steelhead_2007-2020_complied_genotype_data.csv") 
RRSteelhead 
```

The next goal is the remove unwanted columns
```{r}
RRSteelhead
#remove the following columns Genotype_Number, SexID, X37, 
#added a column with the spawn year, the year was taken from the collection date
RRSteelhead$COLLECTION_DATE <- mdy(RRSteelhead$COLLECTION_DATE)
RRSteelhead
RRprep <- RRSteelhead %>% 
  select(-GENOTYPE_NUMBER, -starts_with("SexID"))
RRprep1 <- RRprep %>% 
  mutate(.,SpawnYear = case_when(
    month(RRprep$COLLECTION_DATE) == 1 ~ year(RRprep$COLLECTION_DATE),
    month(RRprep$COLLECTION_DATE) == 2 ~ year(RRprep$COLLECTION_DATE),
    month(RRprep$COLLECTION_DATE) == 3 ~ year(RRprep$COLLECTION_DATE),
    month(RRprep$COLLECTION_DATE) == 4 ~ year(RRprep$COLLECTION_DATE),
    month(RRprep$COLLECTION_DATE) == 5 ~ year(RRprep$COLLECTION_DATE),
    month(RRprep$COLLECTION_DATE) == 12 ~ (year(RRprep$COLLECTION_DATE)+1),
    is.na(RRprep$COLLECTION_DATE) ~ 2013))
#write_csv(RRprep1, "RR_spawn_year_code_check.csv")

RRprep1 %>% select(COLLECTION_DATE, SpawnYear) %>% filter(month(.$COLLECTION_DATE) == 12)
# successful assigned the december spawners into the appropriate spawn year
#only 3 december spawning fish

RRprep1 %>% filter(is.na(SpawnYear)) 
```

Removing loci from analysis
```{r}
#removing the following loci: SH121006-131, Omy-R04944
RRprep1 %>%  
  select(NMFS_DNA_ID, COLLECTION_DATE, SpawnYear,starts_with("SH121006-131"), starts_with("Omy_R04944"))%>%
  filter(is.na(Omy_R04944))
#SH121006-131 was genotyped in 2007 through 2014

RRprep1 %>%  
  select(NMFS_DNA_ID, COLLECTION_DATE, SpawnYear,starts_with("SH121006-131"), starts_with("Omy_R04944"))%>%
  filter(!is.na(Omy_R04944))
#Omy-R04944 was genotyped 2015 through 2020


RRprep2 <- RRprep1 %>% 
  select(-starts_with("SH121006-131"), -starts_with("Omy_R04944")) %>% 
  mutate(Wild = ifelse((STRAIN == "Wild" | grepl("Wild|WILD|wild", SAMPLE_COMMENTS)), yes = "wild", no = NA)) %>% 
  select(NMFS_DNA_ID, BOX_ID, Wild, everything())  
RRprep2$Wild <- replace_na(RRprep2$Wild, "hatchery")
##added a "Wild" column to indicate the fish was thought to be wild (vs hatchery orgin) when spawned
RRprep2 %>%  count(Wild)
```
Next thing to fix is to ensure the metadata is uniform for sex and hatchery and watershed, (ie make "Warm Springs" and "Warm Springs Hatchery" into "Warm_Springs" for all of them)

```{r}
RRprep2 %>% filter(SEX != "Male", SEX != "Female" )
RRprep2 %>% select(SEX) %>% unique()
#there is one problematic sample, M056931, where female is misspelled
RRprep2$SEX <- recode(RRprep2$SEX, "Fmeale" = "Female")  

RRprep2 %>% filter(SEX != "Male", SEX != "Female" )
#misspelling seems to be fixed
  
RRprep2 %>% select(HATCHERY) %>% unique()
RRprep2$HATCHERY <- recode(RRprep2$HATCHERY, "Warm Springs Hatchery" = "Warm_Springs", "Warm Springs" = "Warm_Springs", "Coyote Valley Fish Facility" = "Coyote_Valley", "Coyote Valley Hatchery" = "Coyote_Valley")
RRprep2$WATERSHED <- recode(RRprep2$WATERSHED, "Russian River" = "Russian_River")

RRprep2 %>% select(HATCHERY) %>% unique()
RRprep2 %>% select(SEX) %>%  unique()
RRprep2 %>% select(WATERSHED) %>% unique()
RRprep2 %>% count(WATERSHED)

write_csv(RRprep2, "RR_steelhead_2007-20_compiled_and_clean.csv")
```


The next steps: 1) finding and removing duplicates for a single NMFS_DNA_ID by picking the one with the most data and 2) finding and removing samples with too much missing data (ie "High-missers"), we want to keep the sample with less than 10% failed/missing loci
```{r}
RRprep2 %>% count(SpawnYear)

# first identify which rendition of an individual to keep
rendition_genos <- RRprep2 %>%
  count(NMFS_DNA_ID) %>%
  filter(n != 1) %>%
  semi_join(RRprep2, ., by = "NMFS_DNA_ID") %>%
  group_by(NMFS_DNA_ID) %>%
  mutate(rendition = 1:n()) %>%
  select(rendition, everything()) %>%
  select(-(BOX_ID:Sample_ID), -SpawnYear)

rendition_genos_meta <- RRprep2 %>%
  count(NMFS_DNA_ID) %>%
  filter(n != 1) %>%
  semi_join(RRprep2, ., by = "NMFS_DNA_ID") %>%
  group_by(NMFS_DNA_ID) %>%
  mutate(rendition = 1:n()) %>%
  select(rendition, everything())

#1554 NMFS_IDs with more than 1 rendition (total of 3140)

keepers <- rendition_genos %>%
  gather(key = "locus", value = "allele_int", -rendition, -NMFS_DNA_ID) %>%
  group_by(NMFS_DNA_ID, rendition) %>%
  summarise(num_non_miss = sum(allele_int != 0) / 2) %>%
  arrange(NMFS_DNA_ID, desc(num_non_miss)) %>%
  slice(1)
#keepers has the renditions of the repeated NMFS_IDs with the most data

keepers

# make a wide genos data frame of those individuals
final_dupie_genos <- rendition_genos %>%
  semi_join(keepers, by = c("NMFS_DNA_ID", "rendition")) %>%
  ungroup() %>%
  select(-rendition)

final_dupie_meta <- rendition_genos_meta %>%
  semi_join(keepers, by = c("NMFS_DNA_ID", "rendition")) %>%
  ungroup() %>%
  select(-rendition)
#this gives the list of NMFS_IDs to keep in the two-column format


# now, we want to make a similar wide file of all the rest,
# dropping the ones that were duplicated.
non_duped_genos <- RRprep2 %>%
  anti_join(rendition_genos, by = "NMFS_DNA_ID") %>%
  select(-(BOX_ID:Sample_ID), -SpawnYear)

non_duped_meta <- RRprep2 %>%
  anti_join(rendition_genos_meta, by = "NMFS_DNA_ID")
#this makes a list of all the genotypes that were not duplicated

wide_genos <- bind_rows(non_duped_genos, final_dupie_genos)
wide_meta <- bind_rows(non_duped_meta, final_dupie_meta)
#this binds the non-duplicates with the keepers
#total of 18132 samples after the duplicates are removed

aes(x = ma_sample_id, y = pa_sample_id)
write_rds(wide_meta, "RR_spawner_meta.rds")
wide_meta %>% count(SpawnYear)
```
The next step is to take the "wide_genos" data and remove any "high-missers", which are samples with missing data from more than 10% of the loci. 
```{r}
wide_genos

hi_miss_genos <- wide_genos %>%
  gather(key = "locus", value = "allele_int", -NMFS_DNA_ID) %>%
  group_by(NMFS_DNA_ID) %>% 
  summarise(num_non_miss = sum(allele_int != 0) / 2) %>%
  arrange(num_non_miss) %>%
  filter(num_non_miss < 85)
  
#the highest amount is 94 loci, want to keep the samples with 10% or less missing data, so I want to keep samples with data for 85 loci and higher

#hi_miss_genos is the list of samples to remove from the data, 38 samples to be removed

#final data set should have 18094 samples after removal of hi miss genos

wide_genos_good <- wide_genos %>% 
  anti_join(hi_miss_genos, by = "NMFS_DNA_ID")

wide_meta_good <- wide_meta %>% 
  anti_join(hi_miss_genos, by = "NMFS_DNA_ID")

write_csv(wide_meta_good, "RR_07-20_metadata.csv")
write_csv(wide_genos_good, "RR_07-20_genotypes.csv")
  
wide_genos_good
wide_meta_good
#18094 samples

```

I have now removed the duplicates and removed the samples with more than 10% missing data! I can now get the data ready for HatcheryPedAgree....

Rearranging data for HatcheryPedAgree R package
```{r}
meta <- wide_meta_good %>% 
  select(NMFS_DNA_ID:Sample_ID, SpawnYear) %>% 
  rename(indiv = NMFS_DNA_ID, sex = SEX, spawner_group = COLLECTION_DATE, year = SpawnYear) %>% 
  select(indiv, sex, spawner_group, year, HATCHERY, LENGTH, WATERSHED, Wild) %>% 
  rename(hatchery = HATCHERY, length = LENGTH, watershed = WATERSHED)

write_rds(meta, "RR_steelhead_2007-2020_meta.rds", compress = "xz")
meta %>% filter(is.na(indiv))
geno <- wide_genos_good %>% 
  gather(key = "locus", value = "allele_int", -NMFS_DNA_ID) %>% 
  mutate(locus = str_replace(locus, "_1$", "")) %>% 
  group_by(NMFS_DNA_ID, locus) %>% 
  mutate(gene_copy = 1:2) %>% 
  ungroup() %>% 
  mutate(allele_int = as.integer(allele_int)) %>% 
  rename(indiv = NMFS_DNA_ID)

write_rds(geno, "RR_steelhead_2007-2020_geno.rds", compress = "xz")

meta %>% count(year)
```


