---
title: "RR 2007-2020, Counting the generations with the pedigrees with igraph, Notebook 13"
author: "Anne Beulke"
output: html_notebook
---
using igraph to count the number of generations represented within the dataset
```{r}
#install.packages("igraph")
#install.packages("rgl")
library(igraph)
library(rgl)
library(tidyverse)

Ped_length_age <- read_csv("RR_2007-2020_pedigree_ages_lengths.csv", col_names = T, cols(
  kid = col_character(),
  ma = col_character(),
  pa = col_character(),
  kid_dayofyear = col_character(),
  parent_dayofyear = col_character(),
  parent_sg = col_character(),
  pa_sg = col_character(),
  ma_sg = col_character(),
  kid_sex = col_character(),
  new_kid_sex = col_character(),
  pa_sex = col_character(),
  new_pa_sex = col_character(),
  ma_sex = col_character(),
  new_ma_sex = col_character(),
  SpawnYear = col_double(),
  kid_year = col_double(),
  kid_year2 = col_double(),
  pa_year = col_double(),
  ma_year = col_double(),
  kid_sg = col_character(),
  kid_hatchery = col_character(),
  pa_hatchery = col_character(),
  ma_hatchery = col_character(),
  OffspCollection = col_character(),
  PopName = col_character(),
  FDR = col_double(),
  Pvalue = col_double(),
  LOD = col_double(),
  P.Pr.C_Se_Se = col_double(),
  P.Pr.Max = col_double(),
  MaxP.Pr.Relat = col_character(),
  TotPaNonExc = col_double(),
  TotMaNonExc = col_double(),
  TotUnkNonExc = col_double(),
  TotPairsMendCompat = col_double(),
  TotPairsMendAndLogL = col_double(),
  TotParsMendLoglAndRank = col_double(),
  TotPairsNonExc = col_double(),
  KidMiss = col_double(),
  PaMiss = col_double(),
  MaMiss = col_double(),
  MI.Kid.Pa = col_double(),
  MI.Kid.Ma = col_double(),
  MI.Trio = col_double(),
  MendIncLoci = col_character(),
  kid_age = col_double(),
  new_parent_sg = col_date(format = ""),
  new_kid_sg = col_date(format = ""),
  pa_age = col_double(),
  ma_age = col_double(),
  kid_length = col_double(),
  ma_length = col_double(),
  pa_length = col_double()
))

Ped_length_age
```

Now I want to select only the columns of interest for animal model attempts
```{r}
Pedigree.prep <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, kid, pa, ma) %>%
  arrange(kid_year)
```

```{r}
AnimalModel.prep <- Ped_length_age %>% select(kid, ma, pa, new_kid_sex, kid_age, kid_hatchery, SpawnYear, kid_length)
```

```{r}

Pedigree.prep %>% select(kid) %>% unique() #13067 unique kid ids

teeny <- Pedigree.prep %>%
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}


kid_paths(kiddos[7269])

result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

ggplot(res, aes(x = path_lengths,y =..prop..)) +
  geom_bar()+
  ggtitle("Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")

res %>% filter(path_lengths == 1) %>% unique()  #2910
res %>% filter(path_lengths == 2) %>% unique()  #3298
res %>% filter(path_lengths == 3) %>% unique()  #3994
res %>% filter(path_lengths == 4) %>% unique()  #2527
res %>% filter(path_lengths == 5) %>% unique()  #650
res %>% filter(path_lengths == 6) %>% unique()  #49
res

res %>% select(kids) %>% unique #7269
res %>% unique() #13428
length(founders) #2647 founders (parents only)
length(kiddos) #7269 (kids only, never parents)
```

#Edited code frorm Eric Anderson
## Make a pedigree figure with all of the 5-generation individuals

I think the way we are going to want to do this is to find all the individuals
that are on a 6-generation path.  We will color those in a certain color.  If any
of those individuals have only one parent on the path, then we should have the other parent as 
well, but not grandparents, etc.   We will plot  these with my 'pedvis' package.  This relies on
GraphViz (I think you can do "brew install graphviz" on the terminal to get that.)  
Then, get the package with:
```{r}
#remotes::install_github("eriqande/pedvis")
library(pedvis)
simple_test_ped()
```

### Find all individuals on a 6-generation path

I think we can do this just by 
```{r}
six_path_func <- function(x) {
  z <- all_shortest_paths(graph, from = x, to = founders, mode = "out")$res
  z[sapply(z, length) >= 7]
}

six_path_tibble <- tibble(kids = kiddos) %>%
  mutate(path_lists = map(.x = kids, .f = six_path_func)) %>%
  mutate(plength = map_int(path_lists, length)) %>%
  filter(plength > 0)
  
```
There are 49 paths of length >= 6 generations.  We obviously want to have any individuals explicitly on
such a path in the pedigree (colored gray). But, we also should have any immediate parents of them that are not on the path
as unshaded nodes.  That will
make a nice plot, I think.  It might still be too many individuals to make a nice pedigree, but let's go for it.

For what we need to do, it will suffice to just have a vector of individuals on those paths, so
let's get those:
```{r}
on_path_ids <- lapply(six_path_tibble$path_lists, function(x) unique(names(unlist(x)))) %>%
  unlist() %>%
  unique()

```
That is 162 individuals.  OK.

Now, we just filter our metadata/pedigree table to offspring that are amongst these 162
individuals:
```{r}
fish_for_big_ped <- as_tibble(Ped_length_age) %>%
  filter(kid %in% on_path_ids)
```

### Plot those with Pedvis

Now that we have all those fish, we can prep them for pedvis.

It looks like we just need a Kid Pa Ma data frame:
```{r}
pped <- fish_for_big_ped %>%
  select(kid, pa, ma)

ped_dump <- ped2dot(x = pped, pa = "pa", ma = "ma", kid = "kid",
        ObsNodes = on_path_ids, 
        outf = "6-gen-paths-ped",
        pfactorNodeStyle = "invis",
        pfactorEdgeStyle = "invis"
        )
  
```

The picture we want is in `6-gen-paths-ped.pdf`.


let's try 3 geneartions: 
###
{r, eval=FALSE}
three_path_func <- function(x) {
  z <- all_shortest_paths(graph, from = x, to = founders, mode = "out")$res
  z[sapply(z, length) >= 4]
}

three_path_tibble <- tibble(kids = kiddos) %>%
  mutate(path_lists = map(.x = kids, .f = three_path_func)) %>%
  mutate(plength = map_int(path_lists, length)) %>%
  filter(plength > 0)
on_path_ids <- lapply(three_path_tibble$path_lists, function(x) unique(names(unlist(x)))) %>%
  unlist() %>%
  unique()
fish_for_big_ped <- as_tibble(Ped_length_age) %>%
  filter(kid %in% on_path_ids)
pped <- fish_for_big_ped %>%
  select(kid, pa, ma)

ped_dump <- ped2dot(x = pped, pa = "pa", ma = "ma", kid = "kid",
        ObsNodes = on_path_ids, 
        outf = "3-gen-paths-ped",
        pfactorNodeStyle = "invis",
        pfactorEdgeStyle = "invis"
        )
###



# Going to see how many generations represented in indivdiaul years
# 2013
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(SpawnYear <= 2013) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("2013, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(10633+5209+167)/20629 # 0.7760434

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(5209	+167)/20629 # 0.260604

```


# 2014
```{r}

teeny <- this %>%
  filter(SpawnYear <= 2014) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}


result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

ggplot(res, aes(x = path_lengths,y =..prop..)) +
  geom_bar()+
  ggtitle("2014, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(11162+6529	+443)/22417 # 0.8089396

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(6529	+443)/22417# 0.311014

```



# 2015
```{r}

teeny <- this %>%
  filter(SpawnYear <= 2015) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}


result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

ggplot(res, aes(x = path_lengths,y =..prop..)) +
  geom_bar()+
  ggtitle("2015, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")

res %>% count(path_lengths)

#2 paths for more
(9437	+ 14168	+ 4873+ 129)/32809 # 0.8719254

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(14168+4873+129)/32809# 0.5842909
```


# 2016
```{r}

teeny <- this %>%
  filter(SpawnYear <= 2016) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}


result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

ggplot(res, aes(x = path_lengths,y =..prop..)) +
  geom_bar()+
  ggtitle("2016, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")

res %>% count(path_lengths)

#2 paths for more
(7900	+18370+10332+325)/41166  # 0.8970267

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(18370+10332+325)/41166 # 0.7051207


```



# 2017

```{r}

teeny <- this %>%
  filter(SpawnYear <= 2017) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}


result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

ggplot(res, aes(x = path_lengths,y =..prop..)) +
  geom_bar()+
  ggtitle("2017, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")

res %>% count(path_lengths)

#2 paths for more
(7284+18254+15022+2814+88)/47857   # 0.9081639

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(18254+15022+2814+88)/47857  #0.7559605

```


# Let's look at individual years by hatchery

# Warm Springs, 2013

```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Warm_Springs") %>% 
  filter(SpawnYear <= 2013) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("WS 2013, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4595+4813+167)/12075 # 0.7929607

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(4813+167)/12075 # 0.4124224

```

# Warm Springs 2014
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Warm_Springs") %>% 
  filter(SpawnYear <= 2014) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("WS, 2014, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4173+5655+411)/12835 # 0.7977406

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(5655+411)/12835 # 0.4726139

```

# Warm Springs 2015
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Warm_Springs") %>% 
  filter(SpawnYear <= 2015) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("WS, 2015, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4003+7911+4578+129)/19605 # 0.8477939

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(7911+4578+129)/19605 # 0.6436113

```

# Warm Springs 2016
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Warm_Springs") %>% 
  filter(SpawnYear <= 2016) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("WS, 2016, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4234+6714+9421+305)/24109 # 0.8575221

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(6714+9421+305)/24109 # 0.681903

```

# Warm Springs 2017
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Warm_Springs") %>% 
  filter(SpawnYear <= 2017) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("WS, 2017, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4547+5791+11699+2659+88)/28224 # 0.8781179

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(5791+11699+2659+88)/28224 # 0.7170139

```

#Coyote Valley

#Coyote Valley, 2013
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Coyote_Valley") %>% 
  filter(SpawnYear <= 2013) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("CV, 2013, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(5824+288)/8488 # 0.7200754

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(288)/8488 # 0.03393025

```

# Coyote Valley 2014
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Coyote_Valley") %>% 
  filter(SpawnYear <= 2014) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("CV, 2014, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(6739+746)/9483 # 0.7893072

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(746)/9483 # 0.07866709

```

#Coyote Valley 2015
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Coyote_Valley") %>% 
  filter(SpawnYear <= 2015) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("CV, 2015, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(5886+4424+113)/12370 # 0.8426031

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(4424+113)/12370 # 0.3667745

```

#Coyote Valley 2016
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Coyote_Valley") %>% 
  filter(SpawnYear <= 2016) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("CV, 2016, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4871+8052+493)/15390 # 0.8717349

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(8052+493)/15390 # 0.5552307

```

# Coyote Valley 2017
```{r}
this <- Ped_length_age %>%
  filter(new_ma_sex == "Female") %>%
  filter(new_pa_sex == "Male") %>%
  select(kid_year, SpawnYear, kid, pa, ma, ma_hatchery) %>%
  arrange(kid_year)


teeny <- this %>%
  filter(ma_hatchery == "Coyote_Valley") %>% 
  filter(SpawnYear <= 2017) %>% 
  as_tibble() %>%
  mutate(kid_year = as.character(kid_year))

lumped <- bind_rows(teeny %>%
            select(kid, pa) %>%
            rename(Parent = pa),
          teeny %>%
            select(kid, ma) %>%
            rename(Parent = ma)
)

graph <- lumped %>%
  graph_from_data_frame(., directed = TRUE, vertices = NULL)
  

outdegrees <- degree(graph, mode = "out")
founders <- names(outdegrees)[outdegrees == 0]
indegrees <- degree(graph, mode = "in")
kiddos <- names(indegrees)[indegrees == 0]

boing <- all_shortest_paths(graph, from = kiddos[500], to = founders, mode = "out")

sapply(boing$res, length)

kid_paths <- function(kid) {
  -1 + all_shortest_paths(graph, from = kid, to = founders, mode = "out")$res %>%
    sapply(., length)
}



result <- tibble(kids = kiddos) %>%
  mutate(path_lengths = map(kids, .f = kid_paths))

res <- unnest(result, cols = c(path_lengths))

res %>%
  ggplot() +
  geom_bar(aes(x = path_lengths,y =..prop..))+
  ggtitle("CV, 2017, Generations represented in the dataset" )+
  xlab("Generations")+
  ylab("Proportion of Lineages")


res %>% count(path_lengths)

#2 paths for more
(4320+8952+1897+51)/17324 #0.87855 

#for siblings, need 2 paths from parent to offspring

#for cousins, need 3 paths from parent to offspring

#3paths or more
(8952+1897+51)/17234 #  0.6324707

```






