---
title: "Heritability of Age, Notebook 11"
author: "Anne Beulke"
output: html_notebook
---
The goal is to make a function for getting heritability values

bop = (xp-xo)(⏀p)/(p(xp))
xp and xo from x values in z score caluculation with z equal to ⏀p and ⏀o(written as Yp and Yo below)
p(xp) = ((2π)^(-1/2))e^(-(xp^2)/2)
z score is caluculated from
 z = ( x - μ ) / σ  
x - x vector (elements of x vector)
μ - mean value of x vector
σ - standard deviation of x vector

```{r}
Yp <- 0.001 #number of affected parents out of all parents
Yo <- 0.040 #number of affected offspring from affected parents, out of all offspring from affected parents

#figured out how to get the z score values!!!!!
Xp <- -qnorm(Yp)#3.09
Xo <- -qnorm(Yo)#1.751

Yp; Yo; Xp; Xo

falconer <- function(Yp, Yo){
  Xp <- -qnorm(Yp)
  Xo <- -qnorm(Yo)
  bop <- {((Xp-Xo)*Yp)/(((2*pi)^(-1/2))*exp(-(Xp^2)/2))}
  h2 <- {2*bop}
  both <- list(bop, h2)
  return(both)}
falconer(Yp, Yo)

reich <- function(Yp, Yo){
  Xp <- -qnorm(Yp)
  Xo <- -qnorm(Yo)
  Zw <- (((2*pi)^(-1/2))*exp(-(Xp^2)/2))/Yp
  rop <- (Xp-Xo*(1-(Xp^2-Xo^2)*(1-(Xp/Zw)))^(1/2))/(Zw +(Xo^2)*(Zw-Xp))
  h2 <- 2*rop
  both <- list(rop, h2)
  return(both)}
reich(Yp, Yo)

edwards <- function(Yp, Yo){
  k <- Yo/Yp
  rop <- (0.57*log(k))/(-log(Yp)-0.44*log(k)-0.18)
  h2 <- 2*rop
  both <- list(rop, h2)
  return(both)}
#this equation may not be correct
edwards(Yp, Yo)

```
```{r}
# Standard Error

falconer_se <- function(Yp, Yo, Np, No){
  Xp <- -qnorm(Yp)
  Xo <- -qnorm(Yo)
  p_xp <- {(((2*pi)^(-1/2))*exp((-Xp^2)/2))}
  p_xo <- {(((2*pi)^(-1/2))*exp((-Xo^2)/2))}
  Zo <- (Xp-Xo)
  Zw <- (p_xp/Yp)
  bop <- {Zo/Zw}
  h2 <- {2*bop}
  Vbop <- {(((1/Zw) - (bop*(Zw-Xp)))^2)*((Yp*(1-Yp))/(Np*(p_xp^2))) + ((Yo*(1-Yo))/((Zw^2)*No*(p_xp^2)))}
  h2_se <- {2*sqrt(Vbop)}
  all <- list(bop, h2, Vbop, h2_se)
  return(all)}

falconer_se(0.001, 0.040, 4000, 3000)

  
```

Okay now that i have functions working to compute the three equation, now i need to connect the functions to my data set
```{r}
library(tidyverse)
#list of unique mas and pas from the AgeMatrix
AgeMatrix <- read_csv("RR_2007-2020_age_matrix.csv")
AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  distinct(ma)
#1947 mothers, 3378 fathers, 7486 trios(when exclude parents from 2018)


#Transition Matrix Info
Fathers <- read_csv("RR_2007-2020_transition_matrix_fathers.csv")
Mothers <- read_csv("RR_2007-2020_transition_matrix_mothers.csv")
MidParent <- read_csv("RR_2007-2020_transition_matrix_midparent.csv")

Fathers
Mothers
MidParent
```

```{r}
#2 yr old fathes and sons
#Yp #number of affected parents out of all parents, so #2yr fathers out of all fathers
#Yo #number of affected offspring from affected parents, out of all offspring from affected parents, so number of 2yr old male offspring out of all male offspring from 2 yr old fathers
pa2yr <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(pa_age == 2) %>% 
  filter(new_pa_sex == "Male") %>% 
  distinct(pa)
pas <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_pa_sex == "Male") %>%  
  distinct(pa)

Yp <- nrow(pa2yr)/nrow(pas)
Np <- nrow(pas)
a <- Fathers %>% filter(pa_age == 2, new_pa_sex == "Male", new_kid_sex == "Male", kid_age == 2)
Yo <- a$freq
b <- Fathers %>% filter(pa_age == 2, new_pa_sex == "Male", new_kid_sex == "Male") %>% summarize(sum(n)) %>% rename(n = 'sum(n)')
No <- b$n

F2 <- falconer(Yp, Yo) 
E2 <- edwards(Yp, Yo) 
R2 <- reich(Yp, Yo) 
FS2 <- cbind(Yp,Yo,F2,E2,R2)
FS2

falconer_se(Yp, Yo, Np, No)
```


```{r}
#2 yr mother and daughters
Yp <- nrow(ma2yr)/nrow(mas)
a <- filter(Mothers, ma_age == 2, new_ma_sex == "Female", new_kid_sex == "Female", kid_age == 2)
Yo <- a$freq
Np <- nrow(mas)
b <- Mothers %>% filter(ma_age == 2, new_ma_sex == "Female", new_kid_sex == "Female") %>% summarize(sum(n)) %>% rename(n = 'sum(n)')
No <- b$n

F2 <- falconer(Yp, Yo) 
E2 <- edwards(Yp, Yo) 
R2 <- reich(Yp, Yo) 
MD2 <- cbind(Yp,Yo,F2,E2,R2)
MD2

falconer_se(Yp, Yo, Np, No)
```

#want to do midparent and all offspring!!
```{r}
#need to get mid parent offspring frequencies with the sexes together
MidParent2 <- AgeMatrix %>%
  filter(SpawnYear < 2018) %>% 
  group_by(mid_age, age_pair, kid_age) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))

MidParent2
#write_csv(MidParent2, "RR_2007-2020_transition_matrix_midparent_all_offspring.csv")
```

```{r}
#midparent and all offspring
#2yr olds
# bop = h2, since there are two parents involved and the coeff of relatedness is equal to 1 instead of 1/2

Yp <- nrow(mid2yr)/nrow(mid)
a <- filter(MidParent2, mid_age == 2, kid_age == 2)
Yo <- a$freq
Np <- nrow(mid)
b <- MidParent2 %>% filter(mid_age == 2) %>% summarize(sum(n)) %>% rename(n = 'sum(n)')
No <- b$n

F2 <- falconer(Yp, Yo) 
E2 <- edwards(Yp, Yo) 
R2 <- reich(Yp, Yo) 
MDAll2 <- cbind(Yp,Yo,F2,E2,R2)
MDAll2

falconer_se(Yp, Yo, Np, No)
sqrt(0.0139328)
```

##how about two year old parents, no matter the sex and 2 yr old offspring no  matter the sex
```{r}
#2 yr old fathes and sons
#Yp #number of affected parents out of all parents, so #2yr fathers out of all fathers
#Yo #number of affected offspring from affected parents, out of all offspring from affected parents, so number of 2yr old male offspring out of all male offspring from 2 yr old fathers
pa2yr <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(pa_age == 2) %>% 
  filter(new_pa_sex == "Male") %>%
  distinct(pa)
ma2yr <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(ma_age == 2) %>% 
  filter(new_ma_sex == "Female") %>%
  distinct(ma)
pas <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_pa_sex == "Male") %>%  
  distinct(pa)
mas <- AgeMatrix %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_ma_sex == "Female") %>%  
  distinct(ma)

Yp <- (nrow(pa2yr) + nrow(ma2yr))/(nrow(pas) + nrow(mas))
#stuck here...need a "transition panel of the ages of offspring from 2 yr old parents
Np <- (nrow(pas) + nrow(mas))

EitherParent2 <- AgeMatrix %>%
  filter(SpawnYear < 2018) %>% 
  filter(new_ma_sex == "Female", new_pa_sex == "Male") %>% 
  mutate(EP2yr = case_when(ma_age == 2 ~ "2",
                           pa_age == 2 ~ "2")) %>% 
  group_by(EP2yr, kid_age) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
#write_csv(MidParent2, "RR_2007-2020_transition_matrix_midparent_all_offspring.csv")


a <- EitherParent2 %>% filter(EP2yr == "2", kid_age == 2)
Yo <- a$freq
b <- EitherParent2 %>% filter(EP2yr == "2") %>% summarize(sum(n)) %>% rename(n = 'sum(n)')
No <- b$n

Either2_falconer <- falconer(Yp, Yo) 
Either2_edwards <- edwards(Yp, Yo) 
Either2_reich <- reich(Yp, Yo) 
either <- cbind(Yp,Yo,Either2_falconer, Either2_edwards, Either2_reich)
either

Xp <- -qnorm(Yp)
  Xo <- -qnorm(Yo)
  bop <- {((Xp-Xo)*Yp)/(((2*pi)^(-1/2))*exp(-(Xp^2)/2))}
  h2 <- {2*bop}

  
falconer_se(Yp,Yo,Np,No)
```

