---
title: "RR Steelhead 2007-2020, realtionship between number of offspring and other traits, additional Omy05 analyses, Notebook 7"
output: html_notebook
---

```{r}
library(tidyverse)

Known_Sexes_fish_list <- read_csv("RR_steelhead_07-20_known_sexes.csv", col_names = T, cols(
  indiv = col_character(),
  year = col_character(),
  spawner_group = col_character(),
  hatchery = col_character(),
  length = col_double(),
  watershed = col_character(),
  Wild = col_character(),
  sex = col_character(),
  new_sex = col_character()
  ))
Known_Sexes_fish_list <- Known_Sexes_fish_list %>% 
  replace_with_na(replace = list(length = 203.2)) 

RR_w_Omy <- read_csv("RR_2007-2020_Omy5_meta_pedigree.csv")
```


what proportion of indivdual male and females did not have offspring return and be sampled?
```{r}
Known_Sexes_fish_list %>% count(year)

Potential_parents <- Known_Sexes_fish_list %>% 
  separate(year, c("year1", "year2"), sep = ",") %>% 
  filter(year1 < 2018) %>% 
  filter(!is.na(new_sex))
Potential_parents %>% count( new_sex)
Potential_parents %>% distinct(indiv)
#3757 mas
#9837 pas

Potential_parents %>% 
  filter(new_sex == "Female") %>% 
  inner_join(Known_Sex_Date_Ped, by = c("indiv" = "ma")) %>% 
  distinct(indiv)
#3078 mas with offspring
3078/3757 # have offspring
679/3757 # no offspring as broodstock

Potential_parents %>% 
  filter(new_sex == "Male") %>% 
  inner_join(Known_Sex_Date_Ped, by = c("indiv" = "pa")) %>% 
  distinct(indiv)
#5239 pas with offspring
5239/9837 # have offspring
4598/9837 # no offspring as broodstock

Known_Sexes_fish_list %>% count(new_sex)
```

Any commonalities between the females and males with no offspring returning? 
```{r}

Potential_parents %>% 
  filter(new_sex == "Female") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "ma")) %>%
  count(year1)
679/3757 # no offspring as broodstock

Potential_parents %>% 
  filter(new_sex == "Male") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "pa")) %>% 
  count(year1)
4598/9837 # no offspring as broodstock

#female and males without offspring represent all yeras

Potential_parents %>% 
  filter(new_sex == "Female") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "ma")) %>%
  count(hatchery, year1)

Potential_parents %>% 
  filter(new_sex == "Male") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "pa")) %>%
  count(hatchery, year1)


potential_ma <- Potential_parents %>% 
  filter(new_sex == "Female") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "ma"))
potential_pa <- Potential_parents %>% 
  filter(new_sex == "Male") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "pa"))

potential_ma
potential_pa
```

```{r}
age <- Known_Sex_Date_Ped %>% select(kid, kid_age)

potential_ma_age <- Potential_parents %>% 
  filter(new_sex == "Female") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "ma")) %>%
  left_join(age, by = c("indiv" = "kid")) %>% 
  rename(age =  kid_age)

potential_pa_age <- Potential_parents %>% 
  filter(new_sex == "Male") %>% 
  anti_join(Known_Sex_Date_Ped, by = c("indiv" = "pa")) %>%
  left_join(age, by = c("indiv" = "kid")) %>% 
  rename(age =  kid_age)

potential_ma_none <- potential_ma_age %>% count(age) %>% 
  filter(!is.na(age)) %>% 
  mutate(freq = n/sum(n))

potential_pa_none <- potential_pa_age %>% count(age) %>% 
  filter(!is.na(age)) %>% 
  mutate(freq = n/sum(n))

potential_ma_none
potential_pa_none

```
# compare the freq of ages of fish with and without offspring
```{r}
#parents with offsrping
offspring_pa <- Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_pa_sex == "Male") %>% 
  mutate(pair = paste(ma,pa , sep = "_")) %>% 
  filter(!is.na(pa_age)) %>% 
  distinct(pa, .keep_all = T) %>% 
  count(pa_age) %>% 
  mutate(freq = n/(sum(n)))

offspring_ma <- Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018) %>% 
  filter(new_ma_sex == "Female") %>% 
  mutate(pair = paste(ma,pa , sep = "_")) %>% 
  filter(!is.na(ma_age)) %>% 
  distinct(ma, .keep_all = T) %>% 
  count(ma_age) %>% 
  mutate(freq = n/(sum(n)))
  
offspring_pa
offspring_ma
#no check if significantly different with chi squared test

chisq.test(x = potential_ma_none$n, p = offspring_ma$freq) #sig
chisq.test(x = potential_pa_none$n, p = offspring_pa$freq) #sig

```
# trend with length?
```{r}
potential_ma_age %>% 
  filter(!is.na(length)) %>% #344 with length data
  ggplot()+
  geom_histogram(aes(x = length, y = ..count..))
potential_pa_age %>% 
  filter(!is.na(length)) %>%  #2389 with length data
  ggplot()+
  geom_histogram(aes(x = length, y = ..count..))
#dont see signs of a trend
```

# trend with omy05?
```{r}
omy <- RR_w_Omy %>% select(kid, kid_SH114448_types)
indiv_date <- fishlist_age_spawnday %>% select(indiv, dayofyear)

pot_ma_data <- potential_ma_age %>% 
  left_join(omy, by = c("indiv" = "kid")) %>% 
  left_join(indiv_date, by = "indiv" )

pot_pa_data <- potential_pa_age %>% 
  left_join(omy, by = c("indiv" = "kid")) %>% 
  left_join(indiv_date, by = "indiv" )

pot_ma_freq <- pot_ma_data %>% 
  filter(!is.na(kid_SH114448_types), kid_SH114448_types != "NA_NA") %>% 
  count(kid_SH114448_types) %>% 
  mutate(freq = n/sum(n))

pot_pa_freq <- pot_pa_data %>% 
  filter(!is.na(kid_SH114448_types), kid_SH114448_types != "NA_NA") %>% 
  count(kid_SH114448_types) %>% 
  mutate(freq = n/sum(n))

Kid_Pa_Ma_Age <- read_csv("RR_2007-2020_pedigree_ages.csv")
Kid_Pa_Ma_Age_Pair <- Kid_Pa_Ma_Age %>%  mutate(pair = paste(pa, ma, sep = ":"))

ma_omy_freq <- Kid_Pa_Ma_Age_Pair %>% 
  left_join(omy, by = c("ma" = "kid")) %>% 
  rename(ma_omy = kid_SH114448_types) %>% 
  filter(new_ma_sex == "Female", SpawnYear < 2018) %>% 
  filter(!is.na(ma_omy), ma_omy != "NA_NA") %>% 
  distinct(ma, .keep_all = T) %>% 
  count(ma_omy) %>% 
  mutate(freq = n/sum(n))

pa_omy_freq <- Kid_Pa_Ma_Age_Pair %>% 
  left_join(omy, by = c("pa" = "kid")) %>% 
  rename(pa_omy = kid_SH114448_types) %>% 
  filter(new_pa_sex == "Male", SpawnYear < 2018) %>% 
  filter(!is.na(pa_omy), pa_omy != "NA_NA") %>% 
  distinct(pa, .keep_all = T) %>% 
  count(pa_omy) %>% 
  mutate(freq = n/sum(n))

chisq.test(x = pot_ma_freq$n, p = ma_omy_freq$freq) #sig
chisq.test(x = pot_pa_freq$n, p = pa_omy_freq$freq) #sig

```

# trend with spawn date? 
```{r}
pot_ma_data %>% 
  ggplot()+
  geom_bar(aes(x = dayofyear, y = ..count..))
pot_pa_data %>% 
  ggplot()+
  geom_bar(aes(x = dayofyear, y = ..count..))

Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018, new_ma_sex == "Female") %>% 
  distinct(ma, .keep_all = T) %>% 
  ggplot()+
  geom_bar(aes(x = parent_dayofyear, y = ..count..))

Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018, new_pa_sex == "Male") %>% 
  distinct(pa, .keep_all = T) %>% 
  ggplot()+
  geom_bar(aes(x = parent_dayofyear, y = ..count..))
```
```{r}
pot_ma_data %>% 
  filter(!is.na(dayofyear)) %>% 
  summarize(mean(as.numeric(dayofyear)), sd(as.numeric(dayofyear)))
pot_pa_data %>% 
  filter(!is.na(dayofyear)) %>% 
  summarize(mean(as.numeric(dayofyear)), sd(as.numeric(dayofyear)))

Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018, new_ma_sex == "Female") %>% 
  distinct(ma, .keep_all = T) %>% 
  filter(!is.na(parent_dayofyear)) %>% 
  summarize(mean(as.numeric(parent_dayofyear)), sd(as.numeric(parent_dayofyear)))

Kid_Pa_Ma_Age_Pair %>% 
  filter(SpawnYear < 2018, new_pa_sex == "Male") %>% 
  distinct(pa, .keep_all = T) %>% 
  filter(!is.na(parent_dayofyear)) %>% 
  summarize(mean(as.numeric(parent_dayofyear)), sd(as.numeric(parent_dayofyear)))

#no trend with spawn date
```

#lets look at Omy05 in siblings...
```{r}

sib_combos_x <- RR_w_Omy %>% 
  mutate(ppair = paste(ma, pa, sep = "_")) %>% 
  slice_sample(prop = 1) %>% 
  group_by(ppair, kid_age, new_kid_sex, kid_hatchery, kid_SH114448_types) %>% 
  filter(n() != 1) %>% 
  do(as_data_frame(t(combn(.$kid_dayofyear, m = 2)))) %>% 
  rename(sib_1 = V1, sib_2 = V2) 

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_SH114448_types != "NA_NA") %>% 
  ggplot()+
  geom_count(aes(x = as.numeric(sib_1), y = as.numeric(sib_2), color = as.character(kid_SH114448_types)))+
  coord_fixed(ratio = 1)+
  theme_classic()+
  geom_smooth(method = "lm", aes(x = as.numeric(sib_1), y = as.numeric(sib_2)), color = "#193439")+
  scale_color_manual(values = c("#503143", "#79ad9f", "#9a532b", "#c49b60"))+
  facet_grid(kid_SH114448_types~kid_hatchery)
  

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Coyote_Valley", kid_SH114448_types == "A_A") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Warm_Springs", kid_SH114448_types == "A_A") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Warm_Springs", kid_SH114448_types == "A_R") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Coyote_Valley", kid_SH114448_types == "A_R") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Coyote_Valley", kid_SH114448_types == "R_R") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

sib_combos_x %>% 
  filter(!is.na(new_kid_sex), new_kid_sex == "Female", kid_age == 3, kid_hatchery == "Warm_Springs", kid_SH114448_types == "R_R") %>% 
  lm(as.numeric(sib_2) ~ as.numeric(sib_1), .) %>% 
  summary

```

Want to see if the offspring of AR x AR parents are typically 25% AA, 50% AR and 25% RR
Or, are more AA fish returning than RR fish? If so, this could mean the RR fish are staying in the river as residents

```{r}
omy_stuff <- RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA") %>% 
  count(kid_SH114448_types)

omy_stuff
```



```{r}
omy2 <- omy_stuff %>% 
  mutate(n_exp = ifelse(kid_SH114448_types == "A_R", yes = (sum(n)*0.5), no = (sum(n)*0.25))) %>% 
  rename(n_obs = n)

omy2 %>% 
  ggplot +
  geom_col(aes(x = kid_SH114448_types, y = n_obs), fill = "red", alpha = 0.4) +
  geom_col(aes(x = kid_SH114448_types, y = n_exp), fill = "blue", alpha = 0.4)

obs <- c(296,759,300)
exp <- c(0.25, 0.50, 0.25)
chisq.test(obs, p = exp)
#This means we REJECT the null hypothesis that the obs values = expected values
# So this is support of the alternative hypothesis that the obs and expected are not equal

```
```{r}
# AR parent age distributions
RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(!is.na(ma_age)) %>% 
  count(ma_age) %>% 
  mutate(freq = n/sum(n))

RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(!is.na(pa_age)) %>% 
  count(pa_age) %>% 
  mutate(freq = n/sum(n))
```
```{r}
#offspring age distributions
RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA") %>%
  filter(new_kid_sex == "Female") %>% 
  count(kid_age) %>% 
  mutate(freq = n/sum(n))

RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA") %>%
  filter(new_kid_sex == "Male") %>% 
  count(kid_age) %>% 
  mutate(freq = n/sum(n))

RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA") %>%
  filter(new_kid_sex == "Female") %>% 
  count(kid_SH114448_types, kid_age) %>% 
  group_by(kid_SH114448_types) %>% 
  mutate(freq = n/sum(n))

RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA") %>%
  filter(new_kid_sex == "Male") %>% 
  count(kid_SH114448_types, kid_age) %>% 
  group_by(kid_SH114448_types) %>% 
  mutate(freq = n/sum(n))
```


```{r}
woop <- RR_w_Omy %>% 
  filter(SpawnYear < 2017) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_SH114448_types != "NA_NA")  %>% 
  mutate(mid_parent_age = ((pa_age + ma_age)/2))


woop %>% 
  filter(kid_age == mid_parent_age) %>% 
  ggplot()+
  geom_point(aes(y = as.numeric(kid_dayofyear), x = as.numeric(parent_dayofyear), color = kid_SH114448_types))+
  facet_grid(kid_SH114448_types ~ .)+
  geom_smooth(method = "lm", aes(y = as.numeric(kid_dayofyear), x = as.numeric(parent_dayofyear)))


woop %>% 
  ggplot()+
  geom_histogram(aes(x = kid_age, y = ..count.., color = kid_sex), position = "dodge")+
  facet_grid(kid_hatchery ~ kid_SH114448_types)

woop %>% 
  ggplot()+
  geom_histogram(aes(x = kid_age, y = ..count.., color = kid_sex), position = "dodge")+
  facet_grid(. ~ kid_SH114448_types)
  
m_f_omy <- woop %>% 
  count(kid_sex, kid_SH114448_types) %>% 
  rename(n_obs = n) %>% 
  group_by(kid_sex) %>% 
  mutate(n_exp = ifelse(kid_SH114448_types == "A_R", yes = (sum(n_obs)*0.5), no = (sum(n_obs)*0.25)))

m_f_omy

m_f_omy %>% 
  ggplot +
  geom_col(aes(x = kid_SH114448_types, y = n_obs), fill = "red", alpha = 0.4) +
  geom_col(aes(x = kid_SH114448_types, y = n_exp), fill = NA, color = "blue")+
  facet_wrap(~kid_sex)

#females:
fobs <- c(100,174,66)
exp <- c(0.25, 0.50, 0.25)
chisq.test(fobs, p = exp)
#is sig diff from exp

#males:
mobs <- c(196,585,234)
exp <- c(0.25, 0.50, 0.25)
chisq.test(mobs, p = exp)
#males returning are significantly different than the expected 
  
```

```{r}
m_f_omy2 <- woop %>% 
  count(ma_hatchery, kid_sex, kid_SH114448_types) %>% 
  rename(n_obs = n) %>% 
  group_by(ma_hatchery,kid_sex) %>% 
  mutate(n_exp = ifelse(kid_SH114448_types == "A_R", yes = (sum(n_obs)*0.5), no = (sum(n_obs)*0.25)))

m_f_omy2

m_f_omy2 %>% 
  ggplot +
  geom_col(aes(x = kid_SH114448_types, y = n_obs), fill = "red", alpha = 0.4) +
  geom_col(aes(x = kid_SH114448_types, y = n_exp), fill = "blue", alpha = 0.4)+
  facet_wrap(ma_hatchery~kid_sex)

#CV females:
cvfobs <- c(27, 52,23)
exp <- c(0.25, 0.50, 0.25)
chisq.test(cvfobs, p = exp)
#not sig diff from exp

#WS females:
wsfobs <- c(73,122,43)
exp <- c(0.25, 0.50, 0.25)
chisq.test(wsfobs, p = exp)
#is sig diff from exp

#CV males:
cvmobs <- c(63,178,79)
exp <- c(0.25, 0.50, 0.25)
chisq.test(cvmobs, p = exp)
#not significantly different than the expected 

#WS males:
wsmobs <- c(133, 407, 155)
exp <- c(0.25, 0.50, 0.25)
chisq.test(wsmobs, p = exp)
#is sig diff from exp
  

m_f_omy2 %>% 
  filter(ma_hatchery == "Warm_Springs") %>% 
  ggplot +
  geom_col(aes(x = kid_SH114448_types, y = n_obs), fill = "red") +
  geom_col(aes(x = kid_SH114448_types, y = n_exp), fill = NA, color = "blue")+
  facet_wrap(ma_hatchery~kid_sex)


```

```{r}
RR_w_Omy %>% 
  count(kid_sex, kid_SH114448_types) %>% 
  filter(kid_SH114448_types != "NA_NA") %>% 
  group_by(kid_sex) %>% 
  mutate(freq = n/(sum(n))) %>% 
  ggplot()+
  geom_col(aes(x = kid_SH114448_types, y = freq))+
  facet_wrap(~kid_sex)
```

```{r}
RR_w_Omy %>% 
  mutate(parents = paste(ma, pa, sep = "_"), .after = pa) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_sex == "Female", kid_SH114448_types != "NA_NA") %>% 
  count(parents) %>% 
  arrange(desc(n))

RR_w_Omy %>% 
  mutate(parents = paste(ma, pa, sep = "_"), .after = pa) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(kid_sex == "Male", kid_SH114448_types != "NA_NA") %>% 
  count(parents) %>% 
  arrange(desc(n))
```

```{r}
RR_w_Omy %>% 
  mutate(parents = paste(ma, pa, sep = "_"), .after = pa) %>% 
  filter(ma_SH114448_types == "A_R" & pa_SH114448_types == "A_R") %>% 
  filter(parents == "M036463_M036461", kid_SH114448_types != "NA_NA") %>% 
  count(kid_sex, kid_SH114448_types) %>% 
  group_by(kid_sex) %>% 
  mutate(freq = n/(sum(n)))
```

# want to look at spwan time and omy 5 afer agrouped by age

```{r}
Kid_Pa_Ma_Age_Pair
kid_omy <- RR_w_Omy %>% 
  select(kid, kid_SH114448_types)

ped_omy <- Kid_Pa_Ma_Age_Pair %>% 
  left_join(kid_omy, by = c("kid" = "kid")) %>% 
  rename(kid_omy = kid_SH114448_types) %>% 
  filter(!is.na(kid_omy), kid_omy != "NA_NA", kid_age < 5)
```

#now that the data is ready, looking at the omy5 and spawn date by age group
```{r}
ped_omy %>% 
  group_by(kid_age) %>% 
  count(new_kid_sex, kid_omy)

ped_omy %>% 
  filter(as.numeric(kid_dayofyear) < 200) %>% 
  ggplot()+
  geom_histogram(aes(x= as.numeric(kid_dayofyear), y = ..count.., binwidth = 5)) +
  facet_grid(kid_age ~ kid_omy)

ped_omy %>% 
  filter(as.numeric(kid_dayofyear) < 200) %>% 
  group_by(kid_age, kid_omy) %>% 
  summarize(mean(as.numeric(kid_dayofyear)), sd(as.numeric(kid_dayofyear)))

ped_omy %>% 
  filter(as.numeric(kid_dayofyear) < 200) %>% 
  filter(!is.na(new_kid_sex)) %>% 
  group_by(kid_age,  new_kid_sex, kid_omy) %>% 
  summarize(mean(as.numeric(kid_dayofyear)), sd(as.numeric(kid_dayofyear)))
```

